Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-08 robdml.tex
tex
u

Given a set of data points, a metric gives a distance value between each pair
of them.
%
Deep Metric Learning (DML) aims to learn such a metric between two inputs (\eg,
images) leveraging the representational power of deep neural networks.
%
DML has been extensively studied~\cite{revisiting}, and has a
wide range of applications such
as image retrieval~\cite{imagesim2} and face recognition~\cite{facenet,domainface}.
%
Despite the improvements of this field thanks to the advancement of deep neural
networks, recent research works suggest that DML models are vulnerable to
adversarial attacks, where an imperceptible perturbation could incur unexpected
or covertly manipulated results~\cite{advrank,advorder}.
%
Such vulnerability poses safety, security, and fairness concerns in the
applications of DML.
%
For example, impersonation or recognition evation are possible for DML-based
face-identification system.
%
To counter the attacks (or reducing the vulnerability), it is important to
design defense methods to improve the adversarial robustness of DML models.

\begin{figure}[t]
	\includegraphics[width=1.0\columnwidth]{introplot.pdf}
	\caption{\oo{Performance plot with aligned conditions.
	(1) robustness and r1 under the same training cost;
	(2) r1 and training cost for reaching similar robustness;
	(3) training cost and robustness for reaching similar r1.}}
\end{figure}

% existing methods & problem
Several defense methods for DML are proposed in the
literature~\cite{advrank,robrank}, inspired by the Madry's adversarial training
method~\cite{madry}, as it is kown as one of the most effective defense methods
for deep neural network classifiers.
%
However, it has been noted that
%
(1) the robustness level achieved by the existing methods is still insufficient
to counter the attacks;
%
(2) the direct adoptation of Madry's min-max adversarial training~\cite{madry} will easily
lead to model collapse due to producing very hard sample triplets;
%
(3) the adversarial training procedure is very time-consuming compared to
the training process of a regular DML model, but existing defense methods
are incompatible to acceleration methods like Free Adversarial Training~\cite{freeat}.

% for problem 1
\oo{[1. Hardness Manipulation]}
Triplet hardness still matters in adversarial training for deep metric
learning.

@@

\subsection{Selection of Destination Hardness}

@@

Hardness sorting.

Hardness selection \cref{tab:desth}.

@@

%1. adversarial training speed.

% PGD-7 PGD-16

>>>>

Given a set of data points, a metric gives a distance value between each pair
of them.
%
Deep Metric Learning (DML) aims to learn such a metric between two inputs (\eg,
images) leveraging the representational power of deep neural networks.
%
DML has been extensively studied~\cite{revisiting}, and has a
wide range of applications such
as image retrieval~\cite{imagesim2} and face recognition~\cite{facenet,domainface}.

Despite the improvements of this field thanks to the advancement of deep neural
networks, recent research works suggest that DML models are vulnerable to
adversarial attacks, where an imperceptible perturbation could incur unexpected
or covertly manipulated results~\cite{advrank,advorder}.
%
Such vulnerability poses safety, security, and fairness concerns in the
applications of DML.
%
For example, impersonation or recognition evation are possible for DML-based
face-identification system.
%
To counter the attacks (or reducing the vulnerability), it is important to
design defense methods to improve the adversarial robustness of DML models.

\begin{figure}[t]
	\includegraphics[width=1.0\columnwidth]{introplot.pdf}
	\caption{\oo{Performance plot with aligned conditions.
	(1) robustness and r1 under the same training cost;
	(2) r1 and training cost for reaching similar robustness;
	(3) training cost and robustness for reaching similar r1.}}
	\label{fig:introplot}
\end{figure}

% existing methods & problem
Several defense methods for DML are proposed in the
literature~\cite{advrank,robrank}, inspired by the Madry's adversarial training
method~\cite{madry}, as it is kown as one of the most effective defense methods
for deep neural network classifiers.
%
However, it has been noted that
%
(1) the robustness level achieved by the existing methods is still insufficient
to counter the attacks;
%
(2) the direct adaptation of Madry's min-max adversarial training~\cite{madry} will easily
lead to model collapse due to producing very hard sample triplets;
%
(3) the adversarial training procedure is very time-consuming compared to
the training process of a regular DML model, but existing defense methods
are incompatible to acceleration methods like Free Adversarial Training~\cite{freeat}.

% inspiration.
Recall that a DML model is already prone to collapse with very hard but not
adversarially purturbed sample triplets~\cite{facenet}.
%
To overcome this issue, various triplet sampling strategies, such as
semi-hard~\cite{facenet} and distance~\cite{distance} are proposed to provide
effective negative examples to facilitate metric learning while avoiding model
collapse.
%
In the classical min-max adversarial training, the adversarial example triplet
created by maximizing the triplet loss value can incur model collapse more
easily than unperturbed sample triplets.
%

In this paper, we propose Hardness Manipulation (HM) to create adversarial
examples for increasing their hardness towards that of a harder sample triplet.
%
%
%
Triplet hardness still matters in adversarial training for deep metric
learning.

@@

\subsection{Selection of Source/Destination Hardness}

@@

To figure out an appropriate combination for further experiments, we try
every possible combination and compare their R@1 and ERS performance, as
shown in \cref{tab:desth}.
%
\oo{The table suggests that}

\subsection{Comparison with State-of-the-art}

\oo{experiment, 8 step by default}
Number of PGD Steps $\eta$.
\cref{fig:introplot}

@@
